---
title: "Marker regression results"
author: "Jay Gillenwater"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

## target knits Rmds in their own session, so load libraries here.
source(here::here("packages.R"))
```

```{r load-targets, include=FALSE}
tar_load(Marker_regression_summary)
tar_load(Marker_regression_results)
```

# Overview

I ran a simple marker regression analysis on all the phenotypes in the data set. This performs a simple association test at the site of each marker in the cross. Markers are declared to be associated at some significance level at significance thresholds that were calculated with a permutation test that used 1000 permutations of the data. Below I will present the LOD profiles of this analysis and a table of the significant markers and their positions in the linkage map.  

In the LOD plots below, the 10% significance threshold is indicated by the dashed line, and the 5% threshold is indicated by the solid line.

```{r LodProfiles, echo = FALSE, message = FALSE}

# A function to parse the "location" column. This column has the 
# long and cumbersome names that I used to keep track of the location
# and the trait together. Now, simplify this into a more readible 
# format
extract_location_name <- function(longLocName){
  shortLocName <- ifelse(str_detect(longLocName, "LA_"), "LA", 
                         ifelse(str_detect(longLocName, "SAN_"), "SAN", 
                                ifelse(str_detect(longLocName, "MS_"), "MS", 
                                       ifelse(str_detect(longLocName, "CLA_"), "CLA", 
                                              ifelse(str_detect(longLocName, "SC_"), "SC", "Average")))))
  
  return(shortLocName)
}


# Next, a function to make a set of LOD plots 
# with the 5% and 10% significance thresholds 
# shown with soplid and dashed lines, respectively
make_lod_plot <- function(mr_result, pheno_name){
  
   # Pull aside the results and permutations for clearer code
  perms   <- pluck(mr_result, "perms")
  results <- pluck(mr_result, "result")
  
  # Get the 5% and 10% significance thresholds from the permutations
  perm_summaries <- summary(perms)
  
  for(i in 1:ncol(perm_summaries)){
    
    # Get the significance thresholds for the current scan from the 
    # summary.scanoneperm table
    sig_5_pct  <- perm_summaries[1, i]
    sig_10_pct <- perm_summaries[2, i]
    
    # Get the location name from the column names of the permutation summaries
    Loc <- extract_location_name(colnames(perm_summaries)[[i]])
    
    # Make a title from the location and phenotype name
    title <- paste(Loc, pheno_name, sep = " - ")
    
    # Plot the LOD profile
    plot(results, 
         lodcolumn = i, 
         ylab = "LOD", 
         main = title, 
         ylim = c(0, (sig_5_pct + 0.5)))
    
    # Add a dashed line for the 10% threshold and a solid line for the 5% threshold
    abline(h = sig_5_pct)
    abline(h = sig_10_pct, lty = 2)
    
  }
}

walk2(Marker_regression_results, names(Marker_regression_results), make_lod_plot)

```


Below is a table of the QTL detected at a threshold of alpha = 0.1 within each environment, for each phenotype. 
```{r SigQTL_table, echo = FALSE}

Marker_regression_summary %>% 
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>% 
  kbl(caption = "QTL detected at alpha = 0.1") %>% 
  kable_material(c("striped", "hover"))

```
