---
title: "Workflow notebook"
author: "Jay Gillenwater"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
## target knits Rmds in their own session, so load libraries here.
## source("here::here(packages.R"))
```

```{r load-targets, include=FALSE}
tar_load(c(Read_and_clean, EDA_Plots))

# # If your chunk output is shown in-line, then you'll need to wrap tar_load()
# # like so:
# 
# withr::with_dir(here::here(), {
#   tar_load(c(target_1, target_2, target_3))
# })
# 
# # This is not needed when using tar_make() to render the document.

```

# Overview

This document will be my journal of sorts for the development of the workflow for the analysis of the NC Raleigh x Soja mapping population. I'll try to both document the functions I make and also give results of the analysis as I go along. 

## Phenotype file overview

The phenotype input file is **Pheno_five_locations.xlsx**. This file has measurements for the genotypes split into sheets for each of the five locations they were grown in. The three phenotypes that were measured were percent nitrogen, carbon, and sulfur content. Each sheet has additional information relevant to the measurement process like run date and time, and various standards and quantities involved in the calculation of each elemental content. 

## Phenotype data cleaning
To make analysis easier, I want to get all the phenotype data into one dataframe first. This is done with the **read_clean** function. Briefly, this function reads in each of the location sheets, adds the location to the data, and then selects a subset of columns (the sample name, location, and phenotypes). The plot of each sample is then derived from the sample name, and the rep number in turn from the plot number. A fourth phenotype is then calculated by dividing the percent nitrogen by the percent sulfur (n_s_percent_ratio). Any sample that does not have a code is filtered out at this point. It looks like these samples without a code name were the standards and checks. The code, location, and rep columns are converted to factors before the final dataframe is returned. Here's what this data looks like. 

```{r, DataOverview, echo = FALSE}

glimpse(Read_and_clean)

```

## EDA
I'll check out the distributions of the phenotypes, boxplot first. 

```{r, PhenoBoxplot, echo = FALSE}
EDA_Plots$boxplot
```

While there's a few samples that fall outside the 1.5 IQR, it looks like there's one sample in carbon, nitrogen, and sulfur in the LA environment that is way lower than the other samples. 

```{r PotentialOutlier}

Read_and_clean %>% 
  pivot_longer(tidyselect::contains("percent")) %>% 
  group_by(name) %>% 
  top_n(1, 1/value) %>% 
  arrange(value) %>% 
  head()

```
 It looks like these three low observations all come from the same plot. Beyond the three phenotypes, it looks like this plot also has an abnormally high weight. I'll look at the other measurements for this genotype to see if it is only this plot that has very low phenotypes and very high weight. 
 
```{r OutlierComparison}
Read_and_clean %>% 
  filter(code == "2087") %>% 
  arrange(loc, plot)
```

It looks like it's only this one plot with very low values, I'll remove this plot from the analysis it seems like it was probably due to a measurement error. 


## Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## repository
if(requireNamespace('git2r', quietly = TRUE)) {
  git2r::repository()
} else {
  c(
    system2("git", args = c("log", "--name-status", "-1"), stdout = TRUE),
    system2("git", args = c("remote", "-v"), stdout = TRUE)
  )
}

## session info
sessionInfo()
```

</details>
