---
title: "Progress Overview"
author: "Jay Gillenwater"
date: "1/08/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
library(ggpubr)
```

# Introduction

This document is a summary of the analysis that has already been done and the analysis that is planned for the Raleigh x Soja mapping population. 

# Progress so far

## Data descriptions

### Population structure

The mapping population consists of 151 recombinant inbred line (RIL) soybean genotypes in the $F_4$ generation. Genotypes are identified through numeric codes that range from 1901 to 2105. Codes 2104 and 2105 are the NC Raleigh and PI 424025B (Soja) population parents respectively. 

### Phenotypic distributions

\newpage
```{r phenoHistograms, echo = FALSE}
tar_load(Read_and_clean)

AvgData_byLoc <- Read_and_clean %>% 
  select(code, loc, percent_carbon, percent_nitrogen, percent_sulfur) %>% 
  pivot_longer(cols = c("percent_nitrogen", "percent_carbon", "percent_sulfur"), 
               names_to = "phenotype") %>% 
  group_by(code, loc, phenotype) %>% 
  summarise(avg_pheno = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>%
  pivot_wider(names_from = phenotype, values_from = avg_pheno)
  

parent.genos <- c("2104", "2105")

hist_plot_fn <- function(phe = "percent_nitrogen")
{
  gghistogram(AvgData_byLoc, x = phe, 
              label = "code",
              label.select = parent.genos,
              facet.by = "loc", 
              repel = TRUE, 
              label.rectangle = TRUE, 
              title = phe, 
              xlab = "% composition")
}

phenos <- c("percent_carbon", "percent_nitrogen", "percent_sulfur")

all_histograms <- map(phenos, hist_plot_fn)

ggarrange(plotlist = all_histograms, nrow = 3)
```

### Genotypic data summary

## Linkage mapping

### Data cleaning procedures

#### Phenotypes

#### Genotypes

### Final map statistics

# Planned Analyses

## Mapping

## Summaries




